\appendix
\section{Appendix}

\subsection{Proof of \cref{lemma:boundL}}\label{app:proof}

We introduce several lemmas that we will use to prove
\cref{lemma:boundL}.
Observe that this analysis closely follows that of
\cite{chitra20analyzing} except that we generalize it to
the case that $q\geq p$.

\begin{lemma}[Theorem 1.4 in \cite{vu05spectral}]\label{lemma:spectral}
    There are constants $C$ and $C'$ such that the following hold.
    Let $M_{i,j}$ be an independent random variable with mean
    0, variance at most $\sigma^2$ for $1 \leq i \leq j \leq m$,
    and absolute value bounded by $1$ where $\sigma \geq C' \sqrt{m} \log^2 m$.
    Then almost surely
    \begin{align}
        \lambda(M) \leq 2 \sigma \sqrt{m}
        + C \sqrt{\sigma} m^{1/4} \log m \nonumber
    \end{align}
    where $\lambda(M)$ denotes the spectral norm
    (i.e. maximum singular value) of $M$.
\end{lemma}

\begin{lemma}[Extension of Lemma 4.5 in \cite{chitra20analyzing}]\label{lemma:boundA}
    Let $A$ be the adjacency matrix of a graph drawn from the
    SBM with intra-block probability $p$ and inter-block
    probability $q$. Define $\bar{A} = \E[A]$.
    There exists a universal constant $C$ such that if
    $p \geq C \log^4 n/n$ then with probability 99/100,
    \begin{align}
        ||A - \bar{A}||_2 \leq 3\sqrt{n \Mpq} \nonumber.
    \end{align}
\end{lemma}

\begin{proof}[Proof (Omitted in \cite{chitra20analyzing})]
    Consider the matrix $M=A- \bar{A}$ with $m=2n$.
    Each entry $M_{i,j}$ for $i\neq j$ is a binary variable with
    probability $p$ or $q$ of firing.
    Recall that $M_{i,i}=0$.
    Then $\E[M_{i,j}] = \E[A_{i,j}] - \E[A]_{i,j} = 0$ and
    the variance $\sigma^2$ is less than $\Mpq$.
    The absolute value of each entry is bounded by 1,
    since every entry of $A$ is either 0 or 1,
    and every entry of $\bar{A}$ is between 0 and 1.
    We use that $1 \geq \sqrt{\Mpq} > \sigma$ and
    $\sqrt{\sigma} \geq C' (2n)^{1/4} \log 2n$.
    Then \cref{lemma:spectral} yields
    \begin{align}
        \lambda(A - \bar{A}) &\leq 2 \sqrt{\sigma} \sqrt{2n} +
        C \sqrt{\sigma} (2n)^{1/4} \log 2n \nonumber \\
        &\leq 2 \sqrt{\Mpq 2n} + \frac{C}{C'}\sqrt{\sigma}\sqrt{\sigma}
        \leq 3 \sqrt{n \Mpq} \nonumber
    \end{align}
    for sufficiently large $n$.
    Since $A - \bar{A}$ is a square matrix,
    $||A-\bar{A}|| \leq \lambda(A-\bar{A})$, which completes the proof.
\end{proof}

\begin{lemma}[Bernstein Inequality]\label{lemma:bernstein}
    Let $X_1, \ldots, X_m$ be independent random variables
    with variances $\sigma_1^2, \ldots, \sigma_m^2$ and
    $|X_i| \leq 1$ almost surely for $i \in [m]$.
    Let $X = \sum_{i \in [m]} X_i$, $\mu = \E[X]$,
    and $\sigma^2 = \sum_{i \in [m]} \sigma_i^2$.
    Then the following holds:
    \begin{align}
        \Pr (|X-\mu|>\epsilon) \leq \exp \left( \frac{e^2}
        {2\sigma^2 + \epsilon/3} \right) \nonumber
    \end{align}
\end{lemma}

With \cref{lemma:boundA} and \cref{lemma:bernstein},
we can now prove \cref{lemma:boundL}.


\begin{proof}[Proof of \cref{lemma:boundL}]
    Let $D$ be the degree matrix of $G$ and define
    $\E[D] = \bar{D}$. By the triangle inequality,
    $||L - \bar{L}||_2 \leq ||D - \bar{D}||_2 +
    ||A - \bar{A}||_2$.
    By \cref{lemma:boundA}, $||A - \bar{A}||_2 \leq
    3 \sqrt{n \Mpq}$.
    Additionally, $||D - \bar{D}||_2$ is bounded
    by $\max_{i \in [2n]} |D_{i,i} - \bar{D}_{i,i}|$.
    $D_{i,i}$ is a sum of Bernoulli random variables
    with total variance $\sigma^2$ upper bounded by
    $2 n \Mpq$. It follows from \cref{lemma:bernstein}
    and our assumption $p = \Omega(1/n)$ that
    $|D_{i,i} - \bar{D}_{i,i}| \leq C \sqrt{n \log n \Mpq}$
    with probability $1- 1/200n$ for
    fixed universal constant $C$.
    By a union bound, we have that
    $\max_i | D_{i,i} - \bar{D}_{i,i}| \leq \sqrt{n \log n \Mpq}$
    with probability 99/100.
    A second union bound with the event that
    $||A - \bar{A}||_2 \leq 3 \sqrt{n \Mpq}$ yields the lemma with
    $C' = C+1$.
\end{proof}

\section{Supplemental Material}

The code, data, and LaTeX files are all publicly
available at \href{https://github.com/rtealwitter/Filter-Bubbles}
{github.com/rtealwitter/Filter-Bubbles}.

The two data sets we use are described below:
\begin{itemize}
    \item Twitter is a graph with $548$ nodes and $3638$ edges. Nodes correspond to users who posted tweets about the Delhi legislative assembly elections of 2013, and edges represent user interactions debating that election. 
    \item Reddit is a graph with $556$ nodes and $8969$ edges. Nodes represent users who posted in the r/politics subreddit. Edges correspond to users' posts in different subreddits. In particular, two users are connected by an edge if they have both posted in two subreddits (other than r/politics) during the given time period.
\end{itemize}